% $Log: not supported by cvs2svn $
% Revision 1.2  1996/04/30  20:35:00  emery
% corrected spelling of superseeded to superseded.
%
\begin{latexonly}
\newpage
\end{latexonly}
\subsection{mpl2sdds}
\label{mpl2sdds}

\begin{itemize}
\item {\bf description:} Adds \verb|mpl| data files to an SDDS data set.  \verb|mpl| is a simple data format used by the
\verb|mpl| Toolkit, which is now largely superseded by SDDS and will not be supported in the future. 
\item {\bf example:} 
\begin{flushleft}{\tt
mpl2sdds APS\_s\_betax.out APS\_s\_betay.out -output=APSbetas.sdds
}\end{flushleft}
\item {\bf synopsis:}
\begin{flushleft}{\tt
mpl2sdds {\em mplFile} [{\em mplFile}...] -output={\em SDDSFile} [-erase]
}\end{flushleft}
\item {\bf files:}
Any number of {\em mplFile} arguments may be given.  These name files in {\tt mpl} format, which
has between two and four columns of data.  {\tt sdds2mpl} attempts to add all of the columns from each
\verb|mpl| data file to the data set.  However, a column that has the same name as an existing column
will not be used.  By default, the data in the {\tt mpl} files is added to {\em SDDSFile}, if it
exists already.
\item {\bf switches:}
    \begin{itemize}
    \item {\tt -output={\em SDDSfile}} --- Specifies that data be added to file {\em SDDSfile}. If
        the file does not exist, it is created. 
    \item \verb|-erase| --- Specifies that if {\em SDDSFile} exists already, it should be
        erased prior to adding any data to the data set.  By default, the data in {\em SDDSFilename}
        is retained.
    \end{itemize}
\item {\bf see also:}
    \begin{itemize}
    \item \progref{sdds2mpl}
    \end{itemize}
\item {\bf author:} M. Borland, ANL/APS.
\end{itemize}

