\begin{latexonly}
\newpage
\end{latexonly}
\subsection{sddsquery}
\label{sddsquery}

\begin{itemize}
\item {\bf description:}

\verb|sddsquery| prints a summary of the SDDS header for a data set.  It also prints 
bare lists of names of defined entities, suitable to use with shell
scripts that need to detect the existence of entities in the data set.
Finally, it will create an SDDS file containing information about what is in the
header.

\item {\bf examples:} 
Get information on the contents of a file:
\begin{flushleft}{\tt
sddsquery APS.twi
}\end{flushleft}
Get a list of the column names only:
\begin{flushleft}{\tt
sddsquery APS.twi -columnList
}\end{flushleft}
Get a list of the column names into a shell variable
\begin{flushleft}{\tt
set names = `sddsquery APS.twi -columnList -delimiter=" "`
}\end{flushleft}
Get a list of the column names into an SDDS file
\begin{flushleft}{\tt
sddsquery APS.twi -columnList -sddsOutput=APS.twi.names
}\end{flushleft}

\item {\bf synopsis:} 
\begin{flushleft}{\tt
sddsquery {\em SDDSfilename} [{\em SDDSfilename}...] 
[-sddsOutput[={\em filename}]]
[\{-arrayList | -columnList | -parameterList | -version\}]
[-delimiter={\em delimitingString}] [-appendUnits[=bare]] [-readAll]
}\end{flushleft}

\item {\bf files:} 
The input filenames may name arbitrary SDDS files.  

If {\tt -sddsOutput} is given, the output normally contains one page for each data
class (i.e., array, parameter, and column).  The following
elements are defined:
        \begin{itemize}
        \item Columns (all string type):
                \begin{itemize}
                \item \verb|Name| --- The name of the element.
                \item \verb|Units| --- The units of the data.
                \item \verb|Symbol| --- The symbol for the element.
                \item \verb|Format| --- The format string for the element (e.g., ``\%f'').
                \item \verb|Type| --- The SDDS data type name (e.g., double, float, etc.).
                \item \verb|Description| --- The description for the element.
                \item \verb|Group| --- The group name (for array elements only).
                \end{itemize}
        \item Parameters:
                \begin{itemize}
                \item \verb|Class| --- The SDDS class for the present page.
                \item \verb|Filename| --- The filename being described by the present page.
                \end{itemize}
        \end{itemize}

\item {\bf switches:}

Normal operation of \verb|sddsquery| results in a printout summarizing
the header of each file.  If one of the options is given, however,
this printout will not appear.  Instead, the selected list of names
appears for each file.

    \begin{itemize}
    \item \verb|sddsOutput[={\em filename}]]| --- Requests that output be delivered in SDDS protocol.
        If no {\em filename} is given, the output is delivered to the standard output.
    \item \verb|arrayList| --- In non-SDDS output mode, 
        requests that a list of array names be printed to the standard output,
        one name per line.  In SDDS output mode, requests that only array information be 
        provided.
    \item \verb|columnList| --- In non-SDDS output mode, 
        requests that a list of column names be printed to the standard output,
        one name per line.  In SDDS output mode, requests that only column information be 
        provided.
    \item \verb|parameterList| --- In non-SDDS output mode, 
        requests that a list of parameter names be printed to the standard output,
        one name per line.  In SDDS output mode, requests that only parameter information be 
        provided.
    \item {\tt -version} --- Requests that the SDDS version number of the file be printed to the
        standard output.  Valid in non-SDDS output mode only.
    \item {\tt -delimiter={\em delimitingString}} --- Requests that listed items be separated by the
        given string.  By default, the delimiter is a newline.  Valid in non-SDDS output mode only.
    \item {\tt -appendunits[=bare]} --- Requests that the units of each item be printed directly following the
        item name.  Valid in non-SDDS output mode only.  If the {\tt bare} qualifier is not given,
        then the units are enclosed in parentheses.
    \item {\tt -readAll} --- Forces {\tt sddsquery} to read the entire file.  On some operating systems
        this is necessary when querying compressed files to prevent ``Broken Pipe'' errors.   For large
        files, use of this option will make {\tt sddsquery} slower.
    \end{itemize}
\item {\bf see also:}
    \begin{itemize}
    \item \hyperref{Data for Examples}{Data for Examples (see }{)}{exampleData}
    \end{itemize}
\item {\bf author:} M. Borland, ANL/APS.
\end{itemize}

