\begin{latexonly}
\newpage
\end{latexonly}
\subsection{sddsexpand}
\label{sddsexpand}

\begin{itemize}
\item {\bf description:}

\verb|sddsexpand| reads data pages from an SDDS file and writes a new
SDDS file containing a separate data page for every row in the input
file.  All column definitions from the input file are turned into
parameter definitions in the output file.  In addition all parameter
definitions from the input file are copied to the output file.  Each
output data page contains the values of the columns from a single row
of the input file, along with the values of the parameters from the
same page.  The output file contains no column or array definitions.

\verb|sddsexpand| is essentially the inverse of \verb|sddscollapse|
(except that the column data thrown out in collapsing a file is not
recoverable).

\item {\bf synopsis:} 
\begin{flushleft}{\tt
sddsexpand [{\em inputFile}] [{\em outputFile}] [-pipe[=input][,output]]
}\end{flushleft}

\item {\bf files:} {\em inputFile} is the name of an SDDS data set to
be expanded.  {\em outputFile} is the result.  Note that {\em
outputFile} will not contain any information from any arrays that are
in {\em inputFile}.

\item {\bf switches:} 
\begin{itemize}
        \item {\tt -pipe[=input][,output]} --- The standard SDDS Toolkit pipe option.
        \item {\tt -noWarnings} --- Suppresses warnings about name clashes between
        parameters and columns.  If such a clash occurs, the parameter data is ignored.
\end{itemize}

\item {\bf see also:}
    \begin{itemize}
    \item \progref{sddscollapse}
    \item \progref{sddsbreak}
    \end{itemize}
\item {\bf author:} M. Borland, ANL/APS.
\end{itemize}



