\begin{latexonly} 
\newpage 
\end{latexonly} 
 
\subsection{sddsderef} 
\label{sddsderef} 
 
\begin{itemize} 
\item {\bf description:} 
\verb|sddsderef| allows array and column dereferencing based on constants 
or on data in parameters and columns. 
 
\item {\bf examples:}  
Let \verb|arrayData| be a file containing a string array named  
\verb|MessageText| 
and a column named \verb|MessageIndex| containing integers. 
The integers are indices into the string array.  To create a new column 
called \verb|Message| giving the message text for each index, the 
following command would be used: 
\begin{flushleft}{\tt  
sddsderef arrayData -column=Message,arraySource=MessageText,MessageIndex 
} 
\end{flushleft} 
\item {\bf synopsis:}  
\begin{flushleft}{\tt 
sddsderef [{\em input}] [{\em output}] [-pipe[=input][,output]] 
\\ -column={\em newName},\{columnSource | arraySource\}={\em name},{\em indexName}[,{\em indexName1}...] 
\\ -parameter={\em newName},\{columnSource | arraySource\}={\em name},
\hspace*{10mm}{\em indexName}[,{\em indexName1}...] 
\\ -constant={\em newName},\{columnSource | arraySource\}={\em name},
\hspace*{10mm} {\em indexValue}[,{\em indexValue}...] 
\\ \hspace*{0.1mm} [-outOfBounds=\{exit | delete\}]
}\end{flushleft} 
 
\item {\bf switches:} 
    \begin{itemize} 
   \item {\tt -pipe[=input][,output]} --- The standard SDDS Toolkit pipe option. 
    \item {\tt -column={\em newName},\{columnSource | arraySource\}={\em name},{\em indexName}} \\
        {\tt [,{\em indexName1}...]}
    --- Creates a new column named {\em newName} containing values 
    found by dereferencing the array (or column) {\em name} 
    using the index or indices in the named columns. 
    \item {\tt -parameter={\em newName},\{columnSource | arraySource\}={\em name},{\em indexName}}\\
         {\tt [,{\em indexName1}...]}
    --- Creates a new parameter named {\em newName} containing values 
    found by dereferencing the array (or column) {\em name} 
    using the index or indices in the named parameters.  A new value 
    is thus generated for the parameter for each page. 
    \item {\tt -constant={\em newName},\{columnSource | arraySource\}={\em name},{\em indexValue}}\\
         {\tt [,{\em indexValue}...]}  
    --- Creates a new parameter named {\em newName} containing values 
    found by dereferencing the array (or column) {\em name} using 
    the index or indices given (as explicit values).  Unless the array (or 
    column) varies from page to page, the new parameter will have the 
    same value on each page. 
    \item {\tt -outBounds=\{exit | delete\}} --- Specifies behavior in the event
        that an index value is out of bounds (i.e., less than 0, or greater than
        or equal to the number of array elements or rows).  If {\tt exit} is given,
        the program aborts with an error; this is the default behavior.
        If {\tt delete} is given, then the source row or page for the index
        is omitted from the output.
    \end{itemize} 
\item {\bf author:} M. Borland, ANL/APS. 
\end{itemize} 
