
\begin{latexonly}
\newpage
\end{latexonly}
\subsection{sddsxref}
\label{sddsxref}

\begin{itemize}
\item {\bf description:}
\verb|sddsxref| creates a new data set by adding selected rows from one data set to another data set.
The rows are selected by matching the string or numeric values in a specified column that is present in both of
two pre-existing data sets.  The user may specify which columns of the second data set to take and which to leave.
The user may also transfer parameter and array data.
\item {\bf synopsis:} 
\begin{flushleft}{\tt
sddsxref [-pipe[=input][,output]] [{\em input}] [{\em xRefFile}] [{\em output}]
[-equate={\em columnName} | -match={\em columnName}]
[-reuse[=rows][,page]] [-fillIn]
[-take={\em columnName},...] [-leave={\em columnName},...] 
[-transfer=\{parameter | array\},{\em name}[,{\em name}...]
[-ifis=\{column | parameter | array\},{\em name}[,{\em name}...]
[-ifnot=\{column | parameter | array\},{\em name}[,{\em name}...]
}\end{flushleft}
\item {\bf files:}
        {\em input} is the data set to which data is being added.  {\em xRefFile} is the data set from
        which data is being taken.  Warning: if {\em output} is not given and if {\tt -pipe=out} is not
        specified, {\em input} is overwritten.  For pipe input, the first file listed is taken to be
        {\em xRefFile}.  For pipe input and output, the only file listed is {\em xRefFile}.
\item {\bf switches:}
    \begin{itemize}
    \item {\tt -equate={\em columnName}}, {\tt -match={\em columnName}}--- 
        These options specify the name of a column that exists in both
        {\em input} and {\em xRefFile}.  For \verb|match|, the column must contain string data, while for the 
        \verb|equate| the column must contain numeric data.  For each row in {\em input}, \verb|sddsxref|
        searches {\em xRefFile} to find the first row for which the \verb|match| column is identical or
        the \verb|equate| column is equal, as appropriate.  This
        row is the one from which any data is taken for addition to the row in {\em input}. If
        neither of these options is given, then rows are taken sequentially from {\em xRefFile}
        for each row of {\em input}.
    \item {\tt -reuse[=rows][,page] } --- By default, each row from {\em xRefFile} is matched to one row in {\em input}.
        If {\tt -reuse=rows} is given, each row from {\em xRefFile} may be matched to any number of
        rows in {\em input}.  Also by default, each page of {\em input} is matched with the corresponding
        {\em page} of {\em xRefFile}.  If {\tt -reuse=page} is given, then each page of {\em input} is
        matched anew to the first page of {\em xRefFile}.  The two qualifiers may be given together.
    \item {\tt -fillIn} --- Normally if there is no match in {\em xRefFile} for a row in {\em input},
        that row will not appear in {\em output}.  If {\tt -fillIn} is given, the row will appear,
        but the values for the columns that are being transferred from {\em xRefFile} will be filled
        with zeros and empty strings, as appropriate.
    \item {\tt -take={\em columnName},...}, {\tt -leave={\em columnName},...}--- 
        These options specify which columns of {\em xRefFile} to extract from
        a matching or equal row of {\em xRefFile}
        for addition to a row of {\em input}.  Wildcards may be given in the column names.
        By default, all columns not in {\em input}
        are taken.  If \verb|take| is employed, only the named columns will be taken.  In either case, no
        column specified under \verb|leave| will be taken.  \verb|-leave=*| causes no columns to be
        taken.
    \item {\tt -transfer=\{parameter | array\},{\em name}[,{\em name}...]} --- 
        This option, which may be given multiple times, specifies the names of
        parameters and arrays to be transfered.  Wildcards are not presently supported in this option.
    \item {\tt -ifis=\{column | parameter | array\},{\em name}[,{\em name}...]}, 
        {\tt -ifnot=\{column | parameter | array\},{\em name}[,{\em name}...]} 
        --- These options allow conditional execution.  If any column that is named under
       a \verb|ifis| option is not present, execution aborts.   If any column that is named under 
       a \verb|ifnot| option is present, execution aborts.
    \end{itemize}
\item {\bf sddsmxref} --- {\tt sddsmxref} is a variant of {\tt sddsxref} that permits multiple {\tt -match}
and {\tt -equate} options for more sophisticated cross-referencing.  In other respects, the program is
used just like {\tt sddsmxref}.  {\tt sddsxref} is much faster, however, for single-criterion matching or
equating.
\item {\bf see also:}
    \begin{itemize}
    \item \hyperref{Data for Examples}{Data for Examples (see }{)}{exampleData}
    \item \progref{sddsselect}
    \end{itemize}
\item {\bf author:} M. Borland, L. Emery, H. Shang, R. Soliday ANL/APS.
\end{itemize}

