\begin{latexonly}
\newpage
\end{latexonly}
\subsection{sddssort}
\label{sddssort}

\begin{itemize}
\item {\bf description:} \hspace*{1mm}\\
\verb|sddssort| sorts the tabular data section of a data set by the values in named columns.
Any number of columns may be involved in the sort, and sorting order may be individually specified.
\item {\bf examples:} \\
Sort the APS Twiss file into alphabetical order by element name:
\begin{flushleft}{\tt
sddssort APS.twi APS.twi.sorted -column=ElementName
}\end{flushleft}
Same, but keep only one instance of each row with the same element name:
\begin{flushleft}{\tt
sddssort APS.twi APS.twi.sorted -column=ElementName -unique
}\end{flushleft}
\item {\bf synopsis:} \\
\begin{flushleft}{\tt
sddssort [-pipe=[input][,output]] [{\em SDDSinput}] [{\em SDDSoutput}] 
-column={\em name}[,\{increasing | decreasing\}] [-column...] 
[-parameter=<name>[,{increasing|decreasing}]...]
[-numericHigh] [-nonDominateSort] 
[-unique[=count]] [-noWarnings]
}\end{flushleft}
\item {\tt files:}\\
{\em SDDSinput} is an SDDS file to be sorted.  If it contains multiple data pages, they
are treated separately.  {\em Warning: } if {\em SDDSoutput} is not given and {\tt -pipe=output}  is not
specified, then {\em SDDSinput} will be replaced.
\item {\bf switches:}
    \begin{itemize}
    \item {\tt -pipe=[input][,output]} --- The standard SDDS pipe option.
    \item {\tt -column={\em name}[,\{increasing | decreasing\}]} --- Requests that the column 
        {\em name} be used to order the rows of each tabular
        data section.  Each subsequent \verb|column| request specifies a subsort of the ordering produced
        by the previous requests.  The \verb|increasing| and \verb|decreasing| keywords may be given to
        specify the desired ordering of the (sub)sort, with increasing order being the default.
     \item {\tt -parameter={\em name}[,\{increasing | decreasing\}]} --- similar to column requests, but sort
        the data pages by parameters.
    \item {\tt -unique[=count]} --- Specifies that for any rows that are identical in the sort column values,
        only the first should be included in the output file.  If the {\tt count} qualifier is given,
        then a count of the number of identical rows is supplied in
        a column called IdenticalCount.
    \item {\tt -nonDominateSort} --- perform non-dominated-sort when multiply sorting columns supplied.
        non-dominated-sort only works for numeric columns. 
    \item {\tt -numericHigh} --- works for string sorting which rank the numeric characters higher than 
       other characters in a string comparison. It also ranks numeric character sets with fewer characters 
       below numeric character sets with more characters.
    \item {\tt -noWarnings} --- Suppresses warning messages.
    \end{itemize}
\item {\bf author:} M. Borland, ANL/APS.
\end{itemize}

